Sediment transport capacity varies more in subglacial systems compared to fluvial ones

### 2023-02-28 theory.tex
small updates
Updates from Overleaf

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.9\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Velocity $v_g$ and $v_f$ magnitude are shown by arrow length and marks at the channel boundary. $S_g$ and $S_f$ represent the cross-sectional area in glacial and fluvial channels.  Note that in the fluvial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

@@

  \begin{table}[H]
    \centering
    \caption{Variables, parameters, and constants }
    \begin{tabular}{ l  c  c c }
      Name &Symbol&  Value&Units \\ \hline
      \textbf{Variables}  & & & \\
      Channel hydraulic diameter (glacier, fluvial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
      Width of channel floor (glacial, fluvial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
      Channel cross-sectional area (glacier, fluvial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
      Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      % Gradient of hydraulic potential (glacier) &$\Psi$ & & $\mathrm{Pa\, m^{-1}}$\\
      Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
      Water velocity (glacier, fluvial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
      Shear stress (glacier, fluvial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
      Width-integrated shear-stress (glacier, fluvial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
      Reynolds number &$R_e$& & $\mathrm{(-)}$\\
      Variable &$\Gamma$&&$\mathrm{(-)}$\\
      Variability of a variable ($\Gamma$) &$\chi$& &$\mathrm{(-)}$\\
           &&&\\

@@

  \subsection{Subglacial channel  parameterization}
  \label{sect:sub_mode}
  To parameterize the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water. 
  To accomplish this, a version of hydraulics parameterization presented in \citet{delaney2019} is used.
  The parameterization uses the assumption that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that a channel and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe to describe this relationship \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
  Here, it is assumed that the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$:
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW}
      D_h = \big(s\, f_r\,\rho_w\, \frac{Q_w^{*\,2}}{\Psi^*}\big)^{\frac{1}{5}}~~.
    \end{equation}
  \end{linenomath*}
  % 
  The representative hydraulic gradient is based upon the pressurized flow in Equation~\ref{eq:DW},
  \begin{equation}
    \label{eq:psi}
    \Psi^*= f_f \,  \rho_i \, g (\frac{\partial  z_{sg}}{\partial x} - \frac{\partial z_c}{\partial x}) +  \rho_i \, g \, \frac{\partial z_c}{\partial x},
  \end{equation}
  % 
  \noindent
  where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\frac{\partial z_{sg}}{\partial x}$ is the surface slope of the glacier and $\frac{\partial z_c}{\partial x}$ is the glacier bed slope.

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.9\linewidth]{model_outputs.pdf}
      \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values. The glacial system is in blue and the fluvial system is in green.  Horizontal bars in d, e, and f represent the range of values in the  output. Gray space represents water discharge below the mean value. }
      \label{fig:model_outs}
    \end{figure}
  \end{center}

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.9\linewidth]{multi_run.pdf}
      \caption{Ratio of range of a variable $\Gamma$ for glacial and fluvial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater (less) than $1$ shown in white (gray) are parameter combinations that vary more in a subglacial (fluvial) system. Plots (d-f) show the response of variable $\Gamma$ to the variability in water discharge ($\chi(Q_w)$) in the parameter combination. 
        Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
      \label{fig:range}
    \end{figure}
  \end{center}

@@

  These different characteristics between glacial and fluvial channels show that records of sediment transport downstream of glaciers capture multiple processes together.
  Such processes may compound at glacier margins, where water flow transitions from pressurized subglacial flow to open-channel flow.
  The inconsistent response of subglacial and fluvial systems to water discharge could add  noise to sediment transport signals in glacial systems compared to fluvial ones.
  In addition, variations in sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone. 
  The differences in the systems' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.

@@

\begin{center}
  \begin{figure}[H]
    \includegraphics[width=0.9\linewidth]{multi_run_vars.pdf}
    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; b)  in glacial and fluvial systems. }
    \label{fig:gammas}
  \end{figure}
\end{center}

>>>>

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.65\linewidth]{Cartoon.pdf}
      \caption{Cartoon of the situation with various parameters. Velocity $v_g$ and $v_f$ magnitude are shown by arrow length and marks at the channel boundary. $S_g$ and $S_f$ represent the cross-sectional area in glacial and fluvial channels.  Note that in the fluvial channel parameterization, rectangular channel shapes are implemented (Section~\ref{sect:fluv}).} 
      \label{fig:cartoon}
    \end{figure}
  \end{center}

@@

  \ian{check units on fluvial channel factor}
  \begin{table}[H]
    \centering
    \caption{Variables, parameters, and constants }
    \begin{tabular}{ l  c  c c }
      Name &Symbol&  Value&Units \\ \hline
      \textbf{Variables}  & & & \\
      Channel hydraulic diameter (glacier, fluvial) &  $D_{hg},\,D_{hf}$&  & $\mathrm{m}$     \\
      Width of channel floor (glacial, fluvial) & $w_{cg},w_{cf}$&  & $\mathrm{m}$     \\
      Channel cross-sectional area (glacier, fluvial) &  $S_g, S_f$& & $\mathrm{m^2}$     \\
      Water discharge (instantaneous) & $Q_w$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      Representative water discharge & $Q_{w}^*$& & $\mathrm{m^{3}\,s^{-1}}$ \\
      % Gradient of hydraulic potential (glacier) &$\Psi$ & & $\mathrm{Pa\, m^{-1}}$\\
      Representative hydraulic gradient  &$\Psi^*$ & & $\mathrm{Pa\, m^{-1}}$\\
      Water velocity (glacier, fluvial)  & $v_g,\,v_{f}$& & $\mathrm{m\,s^{-1}}$ \\
      Shear stress (glacier, fluvial) & $\tau_g,\,\tau_f$&& $\mathrm{Pa \, m^{-1}}$ \\
      Width-integrated shear-stress (glacier, fluvial) & $\tau_{tg},\, \tau_{tf}$&& $\mathrm{Pa \, m^{-1}}$ \\
      Reynolds number &$R_e$& & $\mathrm{(-)}$\\
      Variable &$\Gamma$&&$\mathrm{(-)}$\\
      Variability of a variable ($\Gamma$) &$\chi$& &$\mathrm{(-)}$\\
           &&&\\

@@

  \subsection{Subglacial channel  parameterization}
  \label{sect:sub_mode}
  To parameterize the shear stress of water flowing across sediments below a glacier, the subglacial channel parameterization evaluates the channel geometry and the velocity of the flowing water. 
  To accomplish this, the hydraulics parameterization presented in \citet{delaney2019} is used.
  The parameterization uses the assumption that the water is transported through subglacial channels \citep[Figure~\ref{fig:cartoon}; ][]{rothlisberger1972}, and that a channel and geometry will respond to a representative water discharge over a certain time period $Q_{w}^*$, using the Darcy-Weisbach formulation for water-flow through a pipe to describe this relationship \citep[e.g.][]{rothlisberger1972,clarke2003,werder2013}.
  Here, it is assumed that the hydraulic diameter of the subglacial channel, $D_h$, responds to a representative water discharge, $Q_{w}^*$, and a representative gradient of the hydraulic potential, $\Psi^*$:
  \begin{linenomath*}
    \begin{equation}
      \label{eq:DW}
      D_h = \big(s\, f_r\,\rho_w\, \frac{Q_w^{*\,2}}{\Psi^*}\big)^{\frac{1}{5}}~~.
    \end{equation}
  \end{linenomath*}
  % 
  The representative hydraulic gradient is based upon the pressurized flow in Equation~\ref{eq:DW},
  \begin{equation}
    \label{eq:psi}
    \Psi^*= f_f \,  \rho_i \, g (\frac{\partial  z_{sg}}{\partial x} - \frac{\partial z_c}{\partial x}) +  \rho_i \, g \, \frac{\partial z_c}{\partial x},
  \end{equation}
  % 
  \noindent
  where, $f_f$ is the flotation fraction, $\rho_i$ is density of ice $\frac{\partial z_{sg}}{\partial x}$ is the surface slope of the glacier and $\frac{\partial z_c}{\partial x}$ is the glacier bed slope.

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{model_outputs.pdf}
      \caption{Response of channel width (a), hydraulic diameter (b),  Reynolds number  (c),  water velocity (d), shear-stress (e), and width-integrated shear-stress (f)  to different water discharge values. The glacial system is in blue and the fluvial system is in green.  Horizontal bars in d, e, and f represent the range of values in the  output. Gray space represents water discharge below the mean value. }
      \label{fig:model_outs}
    \end{figure}
  \end{center}

@@

  \begin{center}
    \begin{figure}[H]
      \includegraphics[width=0.7\linewidth]{multi_run.pdf}
      \caption{Ratio of range of a variable $\Gamma$ for glacial and fluvial systems for water velocity  (a), shear-stress (b), and  width-integrated shear-stress (c) in response to mean water discharge,
        Range  $\chi$ for each variable is given in Equation~\ref{eq:var}. Values greater (less) than $1$ shown in white (gray) are parameter combinations that vary more in a subglacial (fluvial) system. Plots (d-f) show the response of variable $\Gamma$ to the variability in water discharge ($\chi(Q_w)$) in the parameter combination. 
        Histograms (g-i) show the frequency of relative variations across variables in the parameterization.}
      \label{fig:range}
    \end{figure}
  \end{center}

@@

  These different characteristics between glacial and fluvial channels show that records of sediment transport downstream of glaciers capture multiple fluvial processes together.
  Such processes might be particularly relevant at glacier margins, where water transitions from pressurized subglacial flow to open-channel flow.
  The inconsistent response of sediment transport in subglacial and fluvial systems to changing water discharge could add noise to sediment transport signals between the systems.
  In addition, the subglacial parameterization shows that variations in sediment transport capacity in glacier systems respond to a large number of factors such as channel size, ice thickness, and glacier surface slope, which may react to climate forcing differently than water discharge alone. 
  The differences in fluvial and subglacial channels' response to changing water discharge should be considered when examining sediment transport processes in glacierized catchments.

@@

\begin{center}
  \begin{figure}[H]
    \includegraphics[width=0.7\linewidth]{multi_run_vars.pdf}
    \caption{Variability $\chi$ in velocity ($v_g$, $v_f$; a) shear stress ($\tau_g$, $\tau_f$; b) and width-integrated shear stress ($\tau_{tg}$, $\tau_{tf}$; c)  in glacial and fluvial systems. }
    \label{fig:gammas}
  \end{figure}
\end{center}
